Estimation of carbon stocks in Colombian mangroves and associated uncertainties

### 2016-07-01 Article_mangroves.tex
added changes about new models in article

The selected models showed relatively similar values of R\textsubscript{a}\textsuperscript{2}. In all cases, the independent variables explained less than 47\% of total variability, but between the whole group of models,  they were between the best in terms of predictive power. Likewise, these four candidate models showed the lowest Mean Squared Error (MSE) values and higher values for the F-statistic. Additionally, independent variables were significant, Goldfeld-Quant test showed that candidate models do not present heterocedasticity problems, and according to the Durbin Watson test there was not autocorrelation in the residuals. Cooks distance test did not evidence remote observations for any candidate model.

@@

Even though the evidence ratio between models with ``substantial weight" indicates that model (6) is 1.23 times more likely than  the next-best model (4), we decided not to accept model (6) as the unique model recomended to predict carbon stocks in Colombian mangroves due to  the results from the AIC weights that indicate there is not enough evidence to chose only one model. Instead, we propose the biomass estimation through the use of weighted estimations of each of the three candidate models classified as models with ``substantial weight''.

Our candidate models included as predictive variables bioclimatology and a satellite-derived vegetation index, information that is easily accessible for all mangrove areas in Colombia. Nevertheless, models built with these variables did not show a very high predictive power. On the contrary, when a location variable such as ``site" was included in the models, the predictive power of all of them improved significantly, showing D\textsuperscript{2}\% around 87\%-88\% for all candidate models. Location variable ``CEU" did not contribute to explain of the variability in our data, as neither did the interaction site:CEU (Table \ref{tab:multilevel}).

%Requires the booktabs if the memoir class is not being used
\begin{table}[htbp]
   \centering
  \topcaption {Deviance analysis for candidate multilevel models through maximum likelihood estimation for AGB. When Site and CEU (Coastal Environmental Unit) are location variables;  AIC is the akaike information criterion; D\textsuperscript{2}\% is the Deviance}  % requires the topcapt package
   \begin{tabular}{p{2cm}p{4cm}p{2cm}p{2cm}p{2cm}p{2cm}}%Column formatting, @{} suppresses leading/trailing space
      \toprule
       Model  & Model modifications & AIC & Null deviance &Residual deviance & D\textsuperscript{2}\%\\
      \midrule
     (1) & Original model & 95.32& 29.86 & 19.32 & 35.60 \\
     (1) & Site & 58.47 & 29.86 & 3.58 & 88.00 \\
     (1) & CEU & 96.38 & 29.86 &16.24& 45.63 \\
     (1) & Site:CEU & 58.43 & 29.86 & 3.58 & 88.00 \\     
     (2) & Original model & 95.43 & 29.86 & 20.25 & 32.18 \\
     (2) & Site & 61.05 & 29.86 & 4.01 & 86.57 \\
     (2) & CEU & 99.70 & 29.86 & 18.47 & 38.15 \\
     (2) & Site:CEU & 61.05& 29.86 & 4.01 & 86.57 \\          
     (3) & Original model & 93.56 & 29.86 & 19.35 & 35.21 \\
     (3) & Site & 61.04 & 29.86 & 4.01 & 86.57 \\     
     (3) & CEU & 102.60& 29.86 & 18.89 & 36.53 \\     
     (3) & Site:CEU & 61.04 & 29.86 & 4.01 & 86.57 \\          
     (4) & Original model & 92.36 & 29.86 & 18.79 & 37.09 \\
     (4) & Site & 	61.07 & 29.86 & 4.01 & 86.56 \\
     (4) & CEU & 98.58 & 29.86 & 17.99 & 39.75 \\
     (4) & Site:CEU & 61.07 & 29.86 & 4.01 & 86.56 \\
      \bottomrule
  \end{tabular}
   \label{tab:multilevel}
\end{table}

@@

Despite these limitations, we found that it is possible to develop accurate models using remote sensing information. In our case, models including EVI (enhanced vegetation index) and BIO9 (mean temperature of driest quarter) provided the best statistics in all cases. Some of the models also included the bioclimatic variable BIO16 (precipitation in the wettest quarter), consistent with previous studies that found good predictive power for these variables \citep{CONL:CONL12060, Herz1999, Anaya2009}.

@@

We identified model structure as an additional source of uncertainty in estimating C stocks \citep[cf][]{IPCC2006}, and for this reason we do  not recommend one particular model but a set of four models that should be used together. We recommend to weigh the predictions from each of these four models according to the AIC and calculate prediction uncertainty accordingly. If additional information is available in terms of stratified areas for different sites, we recommend the use of multilevel models that account for spatial heterogeneity.

>>>>

The selected models showed relatively similar values of R\textsubscript{a}\textsuperscript{2}. In all cases, the independent variables explained less than 47\% of total variability, but between the whole group of models,  they were between the best in terms of predictive power. Likewise, these six candidate models showed the lowest Mean Squared Error (MSE) values and higher values for the F-statistic. Additionally, independent variables were significant, Goldfeld-Quant test showed that candidate models do not present heterocedasticity problems, and according to the Durbin Watson test there was not autocorrelation in the residuals. Cooks distance test did not evidence remote observations for any candidate model.

@@

Even though the evidence ratio between models with ``substantial weight" indicates that model (6) is 1.23 times more likely than  the next-best model (4), we decided not to accept model (6) as the unique model recomended to predict carbon stocks in Colombian mangroves due to  the results from the AIC weights that indicate there is not enough evidence to chose only one model. Instead, we propose the biomass estimation through the use of weighted estimations of each of the three candidate models classified as models with ``substantial weight'' (models 6, 4 and 3).

Our candidate models included as predictive variables bioclimatology and a satellite-derived vegetation index, information that is easily accessible for all mangrove areas in Colombia. Nevertheless, models built with these variables did not show a very high predictive power. On the contrary, when a location variable such as ``site" or ``CEU'' was included in the models, the predictive power of all of them improved significantly, showing D\textsuperscript{2}\% around 86\%-87\% for all candidate models in the case of ``site'' and D\textsuperscript{2}\% around 46\%-65\%  in the case of ``CEU'' . The interaction site:CEU  did not contribute to explain of the variability in our data (Table \ref{tab:multilevel}).

%Requires the booktabs if the memoir class is not being used
\begin{table}[htbp]
   \centering
  \topcaption {Deviance analysis for candidate multilevel models through maximum likelihood estimation for AGB. When Site and CEU (Coastal Environmental Unit) are location variables;  AIC is the akaike information criterion; D\textsuperscript{2}\% is the Deviance}  % requires the topcapt package
   \begin{tabular}{p{2cm}p{4cm}p{2cm}p{2cm}p{2cm}p{2cm}}%Column formatting, @{} suppresses leading/trailing space
      \toprule
       Model  & Model modifications & AIC & Null deviance &Residual deviance & D\textsuperscript{2}\%\\
      \midrule
     (1) & Original model & 91.57& 29.75& 17.12 & 44.42\\
     (1) & Site & 60.01 & 29.75 & 3.86 & 87.02 \\
     (1) & CEU & 85.55 & 29.75 &12.06& 59.46 \\
     (1) & Site:CEU & 60.01 & 29.75 & 3.86 & 87.02 \\     
     (2) & Original model & 92.76 & 29.75 & 16.78 & 43.60 \\
     (2) & Site & 61.85 & 29.75 & 3.847 & 87.07\\
     (2) & CEU & 84.18 & 29.75 & 11.09 & 62.17 \\
     (2) & Site:CEU & 61.85 & 29.75 & 3.847 & 87.07 \\          
     (3) & Original model & 85.38 & 29.75 & 14.66 & 50.72 \\
     (3) & Site & 61.16 & 29.75 & 3.97 & 86.64 \\     
     (3) & CEU & 87.05& 29.75 & 12.52 & 57.92 \\     
     (3) & Site:CEU &61.16 & 29.75 & 3.97 & 86.64 \\          
     (4) & Original model & 84.13 & 29.75 & 14.21 & 52.24 \\
     (4) & Site & 62.65& 29.75 & 4.13 & 86.13\\
     (4) & CEU & 79.77 & 29.75 & 10.43 & 64.94 \\
     (4) & Site:CEU & 62.65& 29.75 & 4.13 & 86.13 \\
     (5) & Original model & 90.97 & 29.75 & 17.65 & 40.67 \\
     (5) & Site & 65.15 & 29.75 & 4.62 & 84.48 \\
     (5) & CEU & 94.85 & 29.75 & 15.99 & 46.25 \\
     (5) & Site:CEU & 65.15 & 29.75 & 4.62 & 84.48 \\
     (6) & Original model & 84.50 & 29.75 & 15.08 & 49.31 \\
     (6) & Site & 64.56 & 29.75 & 4.55 & 84.71 \\
     (6) & CEU & 88.05 & 29.75 & 13.49 & 54.66 \\
     (6) & Site:CEU &  64.56 & 29.75 & 4.55 & 84.71 \\
           \bottomrule
  \end{tabular}
   \label{tab:multilevel}
\end{table}

@@

Despite these limitations, we found that it is possible to develop accurate models using remote sensing information. In our case, models including EVI (enhanced vegetation index), Latitude and BIO9 (mean temperature of driest quarter) provided the best statistics in all cases. Some of the models also included the bioclimatic variable BIO16 (precipitation in the wettest quarter), consistent with previous studies that found good predictive power for these variables \citep{CONL:CONL12060, Herz1999, Anaya2009}.

@@

We identified model structure as an additional source of uncertainty in estimating C stocks \citep[cf][]{IPCC2006}, and for this reason we do  not recommend one particular model but a set of three models that should be used together. We recommend to weigh the predictions from each of these three models according to the AIC and calculate prediction uncertainty accordingly. If additional information is available in terms of stratified areas for different sites, we recommend the use of multilevel models that account for spatial heterogeneity.
